\textit{MSF: Modulated Sub-graph Finder}

### 2019-02-28 MSF_Paper_FA.tex
comments FA

\usepackage{subcaption}
\usepackage{hyperref}
\usepackage{pdfpages}
\newcommand{\FA}[1]{\begingroup\color{magenta}#1\endgroup}
\newcommand{\ILH}[1]{\begingroup\color{blue}#1\endgroup}
\newcommand{\TODO}[1]{\begingroup\color{red}#1\endgroup}
%opening
\begin{document}

@@

	Unlike ORA, FCS has no artificial cut-off to define a DEG list. FCS works in
	three steps. First it calculates the gene-level statistics including
	correlation of molecular measurements using differential expression of
	individual genes, ANOVA, t-test and Z-score. In the second step the
	statistics of individual genes in a pathway are transformed to an
	individual pathway-level statistic commonly using Kolmogorov-Smirnov
	statistic, mean or median. Finally the statistical significance of the
	pathway-level statistics is assessed. Although FCS covers some of the
	limitations from ORA, it still ignores the topology of genes in a pathway,
	cross-talk and overlap of the pathways~\cite{Khatri2012,Campos}. Pathway
	topology based methods are similar to FCS except that they consider the
	topology of each gene during the gene-level statistics but still lacks to aim
	to link the different functional pathways~\cite{Khatri2012}.

@@

	As input, information on functional links between genes provided by
	e.g.~KEGG or Reactome and information on the differential expression status
	of single genes resulting from a DGEA, are required. As a result the
	analysis returns sub-graphs and their joint confidence scores, reflecting
	how the perturbation is migrated through the network. Furthermore, the
	entry points of perturbation in the networks and overlap with conventional
	pathway categories are returned. The output is prepared in a directed
	adjacency file, convenient for visualization, e.g., with
	StringApp~\cite{StringApp}, available as a Cytoscape plug-in~\cite{Cyto}.

@@

	In the next step, we check if any of the initial modulated sub-graphs 
	could further be extended. This is done when the sub-graph could not be extended any more to the direct neighboring gene due to a high \textit{p}-value, but the neighboring gene to the direct neighboring gene has a low \textit{p}-value. This was required to connect together low \textit{p}-value gene but incorporating a single high \textit{p}-value gene. If the combined \textit{p}-value of the initial modulated sub-graph and the extension genes is smaller than the \textit{p}-value of the initial sub-graph the extension is accepted. All possible extension paths up to \emph{3} (default 2) genes at
	all nodes in the sub-graph are tested. Again, this step is iteratively repeated until
	no further genes are added to the significant differentially expressed
	sub-graphs. This step bridges small gaps of genes without a clear
	differential signal in the DGEA.\newline

@@

	After detection and extension of the modulated sub-graphs, they are tested
	if combining the different sub-graphs score better than on their own. The merging of the sub-graphs is done by depth first search traversal from one sub-graph to the other sub-graph. If the two sub-graphs merge
	with the connector of at most \emph{3} genes (default 2 gene) and the
	combined \textit{p}-value of the merged sub-graph including the bridging
	genes in between is less than the individual \textit{p}-values of the two
	sub-graphs, the two sub-graphs are merged together to one big modulated
	sub-graph. This step is repeated iteratively until no sub-graphs could be
	merged.\newline

@@

	In a last post processing step \texttt{MSF} identifies the trigger points
	of the modulated sub-graphs. These trigger genes are the sources of the
	sub-graphs with only outgoing edges. These genes can be interpreted as the
	possible entry points of perturbation from where the stimulus causes
	downstream effects. Each individual source is given an impact score, depending on the number of down stream genes effected in the corresponding sub-graph, equipping with a way to choose the most significant genes for unknown phenotypes or biological processes. In the same spirit the most downstream genes of the
	modulated sub-graphs are identified and defined as sinks. Due to loops not all
	sub-graphs are guaranteed to have sources or sinks.

@@

	\texttt{MSF} generates a directed network file as an output, containing complete directed interactions of all modulated sub-graphs identified. This file could be imported into Cytoscape~\cite{Cyto} for visualization. Second output is the source \& sink file. It contains for all the modulated sub-graphs, sources and sinks identified with the source impact scores. The third output, source weightage is provided as a node attribute file. It could be imported into Cytoscape to see the impact score of sources as the size of the node and log-fold change could be used to see how the regulation of different gene hubs in the modulated sub-graphs look.

@@

	MSF requires Java version 8 and JDK 1.8. The few package dependencies are already been added to the release. MSF runs fast on a standard laptop computer and so it has normal system requirements. To run MSF, the user must provide two text files, one containing the DGEA and the second one containing directed interactions in an adjacency format file. Example files and a detailed tutorial to use MSF has been provided on github https://github. com/Modulated-Subgraph-Finder/MSF.

@@

	For the earliest time point at 6~hpi, three large modulated sub-graphs were identified with 42,
	139, and 69 genes.
	The modulated sub-graphs consist predominantly of cytokines, chemokines
	(CXCL10, CCL8, CXCL9, CXCL11, CXCR4, CCR7, CCL4L1, CCL3L1, CCL4, CCL8, CCL20, CCL3, CCL19) and Interleukin genes (IL6, IL27, IL23).
	IFNB1 and IFNA1 were both identified as two of the possible sources
	in the most significantly modulated sub-graph identified with 42 genes. The impact score of IFNB1 is 14.5\% and IFNA1 is 8.7\% for the sub-graph they belong to.
	Most of the sources identified by \texttt{MSF} were
	type~I interferon induced genes (Supplement material).
	At 24~hpi seven modulated sub-graphs were
	identified with four main sub-graphs consisting
	of 61, 222, 130 and 242 genes, others being smaller than 6 genes. Again, IFNB1
	and IFNA1 were identified as two sources out of the total sources with 3.9\% and 1.6\% impact score. For the last time-point 48~hpi,
	six modulated sub-graphs were identified. Three of the sub-graphs were
	less than ten genes and main sub-graphs had 217, 224 and 276 genes. IFNB1 and IFNA1 were identified as sources in
	the most significantly modulated sub-graph with an impact score of 2.8\% and 3.7\% (Supplement material).

@@

	Table~\ref{tab:rawVsHartung} compares the results of \texttt{MSF}, namely
	the number of detected sub-modules and their total genes numbers, to a
	simple analysis of mapping significantly modulated genes from the DGEA to
	the network and joining neighbors to modules. The numbers indicate that
	\texttt{MSF} detects less but larger and meaningful sub-modules, applying its statistical
	test. Furthermore, the dependency of the results from the \textit{p}-value cutoff
	choice is demonstrated for the DGEA, which is avoided for \texttt{MSF}
	altogether. It showcases how applying different cut-offs to the \textit{p}-value of genes from edgeR to the sub-graphs identified by \texttt{MSF} breaks the larger sub-graphs to many smaller unconnected sub-graphs, many of which are single genes.

@@

	Three main modulated sub-graphs identified by \texttt{MSF} at 6~hpi are
	shown in Fig.~\ref{fig:Sub-graph6hpi}. The graphs represent the immediate output of the \texttt{MSF}-analysis,
	visualized by StringApp~\cite{StringApp} in Cytoscape~\cite{Cyto}. Each
	node represents a gene part of a modulated sub-graph, whereby the
	associated colors code the functional annotation deduced from KEGG
	Pathways. The cross-talk between the pathways and also the multiple
	employment of many genes is evident. The flow of information between
	the sensors and effectors in this particular example could easily be deduced.
	In more detail, sub-graph~1 (bottom) shows how the activation of toll-like
	receptor, cytokine, chemokine activating jak-stat and mapk genes, together with TNF leads into apoptosis. The next significant sub-graph (sub-graph~2: top right) reveals
	how information from the Extra-cellular matrix (ECM) receptor, which are
	reported to interact with Ebola glycoprotein (GP)~\cite{Veljkovic}, chemokines, cytokines, and cytosolic DNA sensing are directly or indirectly controlling cell growth, differentiation, proliferation and apoptosis. It suggests that dysregulation of these pathways is responsible for modulation of apoptosis. Eventually, sub-graph~3 (top left)
	demonstrates how INFA1 and INFB1 modulates once more, via only a few
	intermediate steps, the apoptotic response of the cell. On the other side cAMP signaling genes activates platelet genes.

	This display case might advertise with how little effort complex data
	can be interpreted, help to apprehend the dynamics of the underlying
	processes and suggest testable hypothesis and potential points of
	intervention.

@@

	jActiveModules~\cite{jActiveModules} is a plugin in Cytoscape that searches for molecular interaction network to find expression activated sub-networks. The method used to score the expression activated sub-networks is close to the method used in \texttt{MSF}.

	The first time-point of Ebola infection data was analyzed using jActiveModule (Version 3.2.1) to compare the modulated sub-graphs identified by \texttt{MSF} and jActiveModule. The input files were same for both tools, number of modules parameter was set to 5. jActiveModule identified 5 modules, the module with the highest pathScore was selected for comparison with \texttt{MSF} identified modulated sub-graphs. The module consists of a single graph with 314 genes. While \texttt{MSF} identified three directed modulated sub-graphs with 42, 69 and 139 genes. The overlap of the common genes identified between \texttt{MSF} and jActiveModule is shown in Fig~\ref{fig:MSFvsJActiveModule}. It is difficult to state which tool identifies more accurate set of genes as modulated, since there is no golden standard data example to refer. Although jActiveModule identifies single connected module, \texttt{MSF} provides directionality, with the identification of possible perturbation sources of the sub-graphs.

@@

	Gene enrichment analysis was performed using Reactome analyze data
	tool~\cite{Reactome} (version 67). Reactome's over-representation analysis tool tests whether
	certain Reactome pathways are enriched for the list of genes submitted to it. Genes from \texttt{MSF} identified
	sub-graphs for each time-point were analyzed for gene enrichment using this tool.  For
	comparison the DEG results from edgeR for the three time-points were filtered using the cut-off of
	adjusted \textit{p}-value > 0.05. This DEG list was used for gene
	enrichment analysis. The compression of \texttt{MSF} identified sub-graphs gene list and the DEG list analysis is shown in Fig~\ref{fig:msfvsreactome}

	All enriched pathways with a cut-off of \textit{p}-value > 0.05 for \texttt{MSF} and DEG list for the three time-points were selected. The comparison shows most of the pathways known from literature
	to be dis-regulated by Ebola infection are enriched in both the enrichment
	analysis. Toll-Like receptor signaling pathway when interacts with EBOV
	glycoprotein (GP), it triggers the activation of
	cytokines~\cite{Olejnik}. Toll-like receptor pathway is expected to be dis-regulated in the early stage of infection, this pathway was not identified as significantly dis-regulated when \textit{p}-value
	cut off DEG list was analyzed for enrichment. Nine toll-like receptor cascades TLR10, TLR2, TLR3, TLR4, TLR5, TLR7/8, TLR9, TLR1:TLR2 and TLR6:TLR2 were identified as dis-regulated from gene enrichment analysis of \texttt{MSF} identified sub-graph genes, not a single one of these cascade was shown to be dis-regulated in pathway enrichment analysis from DEG cut-off list. Since \texttt{MSF}
	considers the complete DEG results, even the weak signal at the earliest
	time-point was detected; for example Toll-like receptor signaling. While \texttt{MSF} is able to catch weak signals, it does not provide information about the functional relationships among genes like Reactome tool.

@@

	Gene set enrichment analysis (GSEA)~\cite{Subramanian15545} is a method to identify classes of genes or proteins that are over-represented in a large set of genes or proteins. GSEA uses statistical approaches to identify significantly enriched or depleted groups of genes. The complete DEG list from DGEA of the first time-point 6 hpi was analyzed using bioconductor package GSEABase (version 1.44.0). GSEA was able to identify Toll like recpetor, Chemokine signaling pathway, Cytosolic DNA-sensing pathway, Jak-STAT signaling pathway, RIG-I-like receptor signaling pathway and apoptosis as the highest ranked pathways. Although GSEA identified the important pathways for Ebola infection, it did not show how the cross-talk and topology of genes between the pathways identifed. \texttt{MSF} and GSEA uses complete DEG list without any cut-off, is why pathways important for Ebola infection showed up even with weak signal genes in it.

@@

	Classic pathway analysis tools aim to detect in lists of significantly
	deregulated genes enriched associations with pathway genes categorized by
	their biological function and their interactions. Thereby, depending on the
	tool, the internal pathway topology is considered or neglected all
	together. The here presented tool, \texttt{MSF}, employs a different
	approach, by aiming to detect sub-graphs in whole gene regulatory networks
	which are significantly deregulated in a concerted manner. To this end,
	neighboring genes in the user provided network are tested for jointly
	common regulation. Exploiting that each gene's abundance, although not
	independent from its neighbors, is measured repeatedly on its own,
	sensitivity can be increased by our applied \textit{p}-value meta-analysis, namely
	Hartung's method. This potentially enables to call just not significant
	modulated genes based on the DGEA to be convincingly called to be part of a
	deregulated gene group.  Furthermore, it allows to identify connected
	sub-graphs, representing the propagation of gene regulation perturbation in
	the input network. A better understanding of this propagation, especially
	the critical spots such as sensors, effectors, and hubs, facilitates the projection of potential intervention points,
	e.g., for drug development. Since \texttt{MSF} only uses interaction
	information in gene regulation network, but not the functional grouping of
	the genes into functional pathways, it is especially adapted to discover so
	called cross-talk between such pathways.

@@

	\begin{figure*}[p]
		\centering
		\fbox{
			\begin{minipage}{13 cm}
				\includegraphics[width=13cm]{MSF_Robustness.png}
				\caption{Shows the percentage of \texttt{MSF} identified sub-graph genes recall
					rates for the three different time points of
					EBOV infection data for 100 simulation where
					Poisson distributed noise was added to the
					experimental deduced reads per gene
					counts.\label{fig:DEGvsMSF}}
				\label{label}
			\end{minipage}
		}
	\end{figure*}

@@

	\begin{figure*}[p]
		\centering
		\fbox{
			\begin{minipage}{15 cm}
				\includegraphics[width=1.01\linewidth]{ReactomeVsMSF.png}
				\caption{The Upset plot shows the number of shared pathways between \texttt{MSF} identified sub-graph gene list and DEG cut-off list for the 3 time-points. All the different toll-like receptor cascades are shown to be in the bar with 164 shared pathways only between \texttt{MSF} at different time-points.}
				\label{fig:msfvsreactome}
			\end{minipage}
		}
	\end{figure*}

>>>>

\usepackage{subcaption}
\usepackage{hyperref}
\usepackage{pdfpages}
\usepackage[normalem]{ulem}
\newcommand{\FA}[1]{\begingroup\color{magenta}#1\endgroup}
\newcommand{\ILH}[1]{\begingroup\color{blue}#1\endgroup}
\newcommand{\TODO}[1]{\begingroup\color{red}#1\endgroup}
%opening
\begin{document}

@@

	Unlike ORA, FCS has no artificial cut-off to define a DEG list. FCS works in
	three steps. First it calculates the gene-level statistics including
	correlation of molecular measurements using differential expression of
	individual genes, ANOVA, t-test and Z-score. In the second step the
	statistics of individual genes in a pathway are transformed to an
	individual pathway-level statistic commonly using Kolmogorov-Smirnov
	statistic, mean or median. Finally the statistical significance of the
	pathway-level statistics is assessed. Although FCS covers some of the
	limitations from ORA, it still ignores the topology of genes in a pathway,
	cross-talk and overlap of the pathways~\cite{Khatri2012,Campos}. Pathway
	topology based methods are similar to FCS except that they consider the
	topology of each gene during the gene-level statistics but still lacks to aim
	to link the different functional pathways~\cite{Khatri2012}. \FA{mention jActiveModules somewhere in the intro}

@@

	As input, information on functional links between genes provided by
	e.g.~KEGG or Reactome and information on the differential expression status
	of single genes resulting from a DGEA, are required. As a result the
	analysis returns sub-graphs and their joint confidence scores, reflecting
	how the perturbation is migrated through the network. Furthermore, the
	entry points of perturbation in the networks and overlap with conventional
	pathway categories are returned \FA{mention impact score and reliablilty of sources here}. The output is prepared in a directed
	adjacency file, convenient for visualization, e.g., with
	StringApp~\cite{StringApp}, available as a Cytoscape plug-in~\cite{Cyto}.

@@

	In the next step, we check if any of the initial modulated
        sub-graphs could further be extended \FA{beyond the immediate
          neighbouring genes. Instead of testing single genes and
          their compatibility to be added, groups of genes are
          considered.}

        \sout{This is done when the sub-graph could not be extended any more
        to the direct neighboring gene due to a high \textit{p}-value,
        but the neighboring gene to the direct neighboring gene has a
        low \textit{p}-value. This was required to connect together
        low \textit{p}-value gene but incorporating a single high
        \textit{p}-value gene.} If the combined \textit{p}-value of the
        initial modulated sub-graph and the extension genes is smaller
        than the \textit{p}-value of the initial sub-graph the
        extension is accepted. All possible extension paths up to
        \emph{3} (default 2) genes at all nodes in the sub-graph are
        tested. Again, this step is iteratively repeated until no
        further genes are added to the significant differentially
        expressed sub-graphs. This step bridges small gaps of genes
        without a clear differential signal in the DGEA.\newline

@@

        \FA{After detection and extension of the modulated sub-graphs,
          each pair of so far identified sub-graphs is tested if its
          combination scores better than each on its own.}
        \sout{After detection and extension of the modulated
          sub-graphs, they are tested if combining the different
          sub-graphs score better than on their own.} The merging of
        the sub-graphs is done by depth first search traversal from
        one sub-graph to the other sub-graph. If the two sub-graphs
        merge with the connector of at most \emph{3} genes (default 2
        gene) and the combined \textit{p}-value of the merged
        sub-graph including the bridging genes in between is less than
        the individual \textit{p}-values of the two sub-graphs, the
        two sub-graphs are merged together to one big\FA{ger}
        modulated sub-graph. This step is repeated iteratively until
        no sub-graphs \sout{could be merged}\FA{can be merged
          anymore}.\newline

@@

	In a last post processing step \texttt{MSF} identifies the
        trigger points of the modulated sub-graphs. These trigger
        genes are the sources of the sub-graphs with only outgoing
        edges. These genes can be interpreted as the possible entry
        points of perturbation from where the stimulus causes
        downstream effects. Each individual source is given an impact
        score, \FA{expressing the relative number of downstream genes
          within the corresponding sub-graph directly connected by
          directed links. This score can be interpreted as an upper
          limit of how much of the sub-graph's perturbation could have
          been introduced by the respective source, and thereby could
          be helpful to prioritize different identified sources for
          larger sub-graphs.}\sout{depending on the number of down
          stream genes effected in the corresponding sub-graph,
          equipping with a way to choose the most significant genes
          for unknown phenotypes or biological processes.} In the same
        spirit \FA{as sources}, the most downstream genes of the
        modulated sub-graphs are identified and defined as sinks. Due
        to loops not all sub-graphs are guaranteed to have sources or
        sinks.

        \TODO{?what about source reliability?}

@@

	\texttt{MSF} generates a directed network file as an output,
        containing complete directed interactions of all modulated
        sub-graphs identified. This file could be imported into
        Cytoscape~\cite{Cyto} for visualization. \sout{Second output
          is the source \& sink file. It contains} \FA{Additionally, a
          file containing details on all sources and sinks for all
          modulated sub-graphs is reported. Furthermore, for
          facilitated visualisation in Cytoscape, a node attribute
          file is provided, containing the source weightage and the
          log-fold chances of all considered genes}. \sout{The third output,
        source weightage is provided as a node attribute file. It
        could be imported into Cytoscape to see the impact score of
        sources as the size of the node and log-fold change could be
        used to see how the regulation of different gene hubs in the
        modulated sub-graphs look.}

@@

	MSF requires Java version 8 and JDK 1.8. The few package
        dependencies are already been added to the release. MSF runs
        fast on a standard laptop computer and so it has normal system
        requirements \TODO{What is fast and what is normal?}. To run
        MSF, the user must provide two text files, one containing the
        DGEA \TODO{DESeq2 edgR?} and the second one containing
        directed interactions in an adjacency format file. Example
        files and a detailed tutorial to use MSF has been provided on
        github https://github. com/Modulated-Subgraph-Finder/MSF.

@@

	For the earliest time point at 6~hpi, three large modulated
        sub-graphs were identified with 42, 139, and 69 genes.  The
        modulated sub-graphs consist predominantly of cytokines,
        chemokines (CXCL10, CCL8, CXCL9, CXCL11, CXCR4, CCR7, CCL4L1,
        CCL3L1, CCL4, CCL8, CCL20, CCL3, CCL19) and Interleukin genes
        (IL6, IL27, IL23).  IFNB1 and IFNA1 were both identified as
        two of the possible sources in the most significantly
        modulated sub-graph identified with 42 genes. The impact score
        of IFNB1 is 14.5\% and IFNA1 is 8.7\% for the sub-graph they
        belong to.  Most of the sources identified by \texttt{MSF}
        were type~I interferon induced genes (Supplement material).
        At 24~hpi seven modulated sub-graphs were identified with four
        main sub-graphs consisting of 61, 222, 130 and 242 genes,
        others being smaller than 6 genes. Again, IFNB1 and IFNA1 were
        identified as two sources out of the total sources with 3.9\%
        and 1.6\% impact score. For the last time-point 48~hpi, six
        modulated sub-graphs were identified. Three of the sub-graphs
        were less than ten genes and main sub-graphs had 217, 224 and
        276 genes. IFNB1 and IFNA1 were identified as sources in the
        most significantly modulated sub-graph with an impact score of
        2.8\% and 3.7\% (Supplement material). \TODO{was the mentioned
          impact score the highest of all sources? How many sources
          are there?}

@@

	Table~\ref{tab:rawVsHartung} compares the results of
        \texttt{MSF}, namely the number of detected sub-modules and
        their total genes numbers, to a simple analysis of mapping
        significantly modulated genes from the DGEA to the network and
        joining neighbors to modules. The numbers indicate that
        \texttt{MSF} detects less but larger and meaningful \TODO{?}
        sub-modules, applying its statistical test. Furthermore, the
        dependency of the results from the \textit{p}-value cutoff
        choice is demonstrated for the DGEA, which is avoided for
        \texttt{MSF} altogether. It showcases how applying different
        cut-offs to the \textit{p}-value of genes from edgeR to the
        sub-graphs identified by \texttt{MSF} breaks the larger
        sub-graphs to many smaller unconnected sub-graphs, many of
        which are single genes.

@@

	Three main modulated sub-graphs identified by \texttt{MSF} at
        6~hpi are shown in Fig.~\ref{fig:Sub-graph6hpi}. The graphs
        represent the immediate output of the \texttt{MSF}-analysis,
        visualized by StringApp~\cite{StringApp} in
        Cytoscape~\cite{Cyto}. Each node represents a gene part of a
        modulated sub-graph, whereby the associated colors code the
        functional annotation deduced from KEGG Pathways. The
        cross-talk between the pathways and also the multiple
        employment of many genes is evident. The flow of information
        between the sensors and effectors can be perceived given the
        directionality of each interaction, indicated by arrows
        \sout{in this particular example could easily be deduced}.  In
        more detail, sub-graph~1 (bottom) shows how the activation of
        toll-like receptor, cytokine, chemokine activating jak-stat
        and mapk genes, together with TNF leads into apoptosis. The
        next significant sub-graph (sub-graph~2: top right) reveals
        how information from the Extra-cellular matrix (ECM) receptor,
        which are reported to interact with Ebola glycoprotein
        (GP)~\cite{Veljkovic}, chemokines, cytokines, and cytosolic
        DNA sensing are directly or indirectly controlling cell
        growth, differentiation, proliferation and apoptosis. It
        suggests that dysregulation of these pathways is responsible
        for modulation of apoptosis. Eventually, sub-graph~3 (top
        left) demonstrates how INFA1 and INFB1 modulates once more,
        via only a few intermediate steps, the apoptotic response of
        the cell. On the other side cAMP signaling genes activates
        platelet genes.

	This display case might advertise with how little effort
        complex data can be \FA{processed and prepared for
          interpretation by the domain expert, to } \sout{interpreted,
          help to} apprehend the dynamics of the underlying processes
        and suggest testable hypothesis and potential points of
        intervention.

@@

	jActiveModules~\cite{jActiveModules} \TODO{citation missing in
          bib} is a plugin in Cytoscape that searches for molecular
        interaction network to find expression activated
        sub-networks. The method used to score the expression
        activated sub-networks is close to the method used in
        \texttt{MSF}. \TODO{what are the differences?}

	The first time-point of Ebola infection data was analyzed
        using jActiveModule (Version 3.2.1) to compare the modulated
        sub-graphs identified by \texttt{MSF} and jActiveModule. The
        input files were same for both tools, number of modules
        parameter was set to 5 \TODO{what does this parameter
          do?}. jActiveModule identified 5 modules, the module with
        the highest pathScore was selected for comparison with
        \texttt{MSF} identified modulated sub-graphs \TODO{why only 1
          out of 5?}. The module consists of a single graph with 314
        genes. While \texttt{MSF} identified three directed modulated
        sub-graphs with 42, 69 and 139 genes. The overlap of the
        common genes identified between \texttt{MSF} and jActiveModule
        is shown in Fig~\ref{fig:MSFvsJActiveModule}. \sout{It is
          difficult to state which tool identifies more accurate set
          of genes as modulated, since there is no golden standard
          data example to refer.} Although jActiveModule identifies
        single connected module, \texttt{MSF} provides directionality,
        with the identification of possible perturbation sources of
        the sub-graphs. \TODO{find reasons why jActiveModules gives so
          much bigger moduls by comparing the details of the method!}

@@

	Gene enrichment analysis was performed using Reactome analyze
        data tool~\cite{Reactome} (version 67) \TODO{on which
          data?}. Reactome's over-representation analysis tool tests
        whether certain Reactome pathways are enriched for the list of
        genes submitted to it. Genes from \texttt{MSF} identified
        sub-graphs for each time-point were analyzed for gene
        enrichment using this tool.  For comparison the DEG results
        from edgeR for the three time-points were filtered using the
        cut-off of adjusted \textit{p}-value \sout{>}\FA{<} 0.05. This
        DEG list was used for gene enrichment analysis. The
        compression of \texttt{MSF} identified sub-graphs gene list
        and the DEG list analysis is shown in
        Fig~\ref{fig:msfvsreactome}

	All enriched pathways with a cut-off of \textit{p}-value
        \sout{>}\FA{<} 0.05 0.05 for \texttt{MSF} and DEG list for the
        three time-points were selected. The comparison shows most of
        the pathways known from literature to be dis-regulated by
        Ebola infection are enriched in both the enrichment
        analysis. Toll-Like receptor signaling pathway when interacts
        with EBOV glycoprotein (GP), it triggers the activation of
        cytokines~\cite{Olejnik}. Toll-like receptor pathway is
        expected to be dis-regulated in the early stage of infection,
        this pathway was not identified as significantly dis-regulated
        when \textit{p}-value cut off DEG list was analyzed for
        enrichment. Nine toll-like receptor cascades TLR10, TLR2,
        TLR3, TLR4, TLR5, TLR7/8, TLR9, TLR1:TLR2 and TLR6:TLR2 were
        identified as dis-regulated from gene enrichment analysis of
        \texttt{MSF} identified sub-graph genes, not a single one of
        these cascade was shown to be dis-regulated in pathway
        enrichment analysis from DEG cut-off list. Since \texttt{MSF}
        considers the complete DEG results, even the weak signal at
        the earliest time-point was detected; for example Toll-like
        receptor signaling. While \texttt{MSF} is able to catch weak
        signals, it does not provide information about the functional
        relationships among genes like Reactome tool.

@@

	Gene set enrichment analysis (GSEA)~\cite{Subramanian15545}
        \TODO{reference is missing in bib} is a method to identify
        classes of genes or proteins that are over-represented in a
        large set of genes or proteins. GSEA uses statistical
        approaches to identify significantly enriched or depleted
        groups of genes. The complete DEG list from DGEA of the first
        time-point 6 hpi was analyzed using bioconductor package
        GSEABase (version 1.44.0). GSEA was able to identify
        Toll\FA{-}like recpetor, chemokine signaling pathway,
        cytosolic DNA-sensing pathway, Jak-STAT signaling pathway,
        RIG-I-like receptor signaling pathway and apoptosis as the
        highest ranked pathways. Although GSEA identified the
        important pathways for Ebola infection, it did not show how
        the cross-talk and topology of genes between the pathways
        identifed. \texttt{MSF} and GSEA uses complete DEG list
        without any cut-off, is why pathways important for Ebola
        infection showed up even with weak signal genes in it.

@@

	Classic pathway analysis tools aim to detect in lists of
        significantly deregulated genes enriched associations with
        pathway genes categorized by their biological function and
        their interactions. Depending on the tool, the internal
        pathway topology is considered or neglected all together. The
        here presented tool, \texttt{MSF}, employs a different
        approach, by aiming to detect sub-graphs in whole gene
        regulatory networks which are significantly deregulated in a
        concerted manner. To this end, neighboring genes in the user
        provided network are tested for jointly common
        regulation. Exploiting that each gene's abundance, although
        not independent from its neighbors, is measured repeatedly on
        its own, sensitivity can be increased by our applied
        \textit{p}-value meta-analysis, namely Hartung's method. This
        potentially enables to call just not significant modulated
        genes based on the DGEA to be convincingly called to be part
        of a deregulated gene group.  Furthermore, it allows to
        identify connected sub-graphs, representing the propagation of
        gene regulation perturbation in the input network. A better
        understanding of this propagation, especially the critical
        spots such as sensors, effectors, and hubs, facilitates the
        projection of potential intervention points, e.g., for drug
        development. Since \texttt{MSF} only uses interaction
        information in gene regulation network, but not the functional
        grouping of the genes into functional pathways, it is
        especially adapted to discover so called cross-talk between
        such pathways.

@@

	\begin{figure*}[p]
		\centering
		\fbox{
			\begin{minipage}{13 cm}
				\includegraphics[width=13cm]{MSF_Robustness.png}
				\caption{\FA{Recall rates for genes in
                                  }\sout{Shows the percentage of}
                                  \texttt{MSF} identified sub-graph
                                  \sout{genes recall rates} for the
                                  three different time points of EBOV
                                  infection data for 100
                                  simulation\FA{s} where Poisson
                                  distributed noise was added to the
                                  experimental deduced reads per gene
                                  counts.\label{fig:DEGvsMSF}}
				\label{label}
			\end{minipage}
		}
	\end{figure*}

@@

	\begin{figure*}[p]
		\centering
		\fbox{
			\begin{minipage}{15 cm}
				\includegraphics[width=1.01\linewidth]{ReactomeVsMSF.png}
				\caption{The Upset plot shows the
                                  number of shared pathways between
                                  \texttt{MSF} identified sub-graph
                                  gene list and DEG cut-off list for
                                  the three time-points. All
                                  \sout{the} \FA{XXX} different
                                  toll-like receptor cascades are
                                  \sout{shown to be} in the \FA{set
                                    of}\sout{bar with} 164 shared
                                  pathways only between \texttt{MSF}
                                  at different time-points.}
				\label{fig:msfvsreactome}
			\end{minipage}
		}
	\end{figure*}
